---
title: "BIOL3100_Final_Project"
author: "Jessica Richards"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, echo=FALSE, include=FALSE}
library(tidyverse)

```

# This is my data about trees

### Source
link to original authors:
https://doi.org/10.5061/dryad.xd2547dpw 

citation:
Wood, K., Kobe, R., Ibáñez, I., et al. (2023). Tree seedling functional traits
mediate plant-soil feedback survival responses across a gradient of light
availability. Dryad. https://doi.org/10.5061/dryad.xd2547dpw

funded by: National Science Foundation: NSF DEB 145732, Michigan State University,
and Alma College

### Methods/What this Data is
This experiment was a factorial blocked design field experiment. 4 tree species 7 soil sources and a variety of forest understory light levels were used. There were a total of 3,024 seedlings. Survival was measured twice a week and seedlings were randomly selected each time to measure mycorrhizal colonization, phenolics, lignin, and NSC at 3 weeks. Cox proportional hazards survival models were used to evaluate survival and linear mixed effects models were used to test how light availability and soil source influence traits. 




## Original Data {.tabset}


### Output

Preview of original Data

```{r original data output, echo=FALSE, include=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
head(tree_dat)

```

### R code

Preview of original Data

```{r original data r code, echo=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
head(tree_dat)

```


# Cleaned Data

### Columns cleaned data 
- Species - 4 species
- Light_ISF vs Light_Cat, Light_ISF
- Core
- Soil
- Adult
- Conspecific
- Myco and SoilMyco
- PlantDate
- AMF vs EMF
- Phenolics
- Lignin
- NSC
- Census vs Time
- Event vs Harvest vs Alive

### what do I want in cleaned data set 
- keep "No"    "Plot"    "Subplot"   "Species"  
- only keep "Light_Cat"
- maybe keep "Core"?
- have only one Soil column with conspecific sterile, species, etc. 
- maybe keep conspecific?
- idk what myco and soilmyco are
- maybe keep adult?
- probably not plantdate unless relevant
- Figure out what AMF and EMF are
- Keep phenolics, lignin, and nsc
- keep time
- Change Event, Harvest, and Alive into one column (well really two)




## Cleaned Data {.tabset}

### Plots

cleaned data has less columns, better for plotting/modeling
```{r clean data head, echo=FALSE, include=TRUE}



```

### Tables

cleaned data has less columns, better for plotting/modeling
```{r clean data r code, echo=TRUE}



```




# Plotting Data

# Modelling Data

# Conclusions


## Results {.tabset}

### Plots

We show a scatter plot in this section.

```{r, fig.dim=c(5, 3)}
par(mar = c(4, 4, .5, .1))
plot(mpg ~ hp, data = mtcars, pch = 19)
```

### Tables

We show the data in this tab.

```{r}
head(mtcars)
```


