---
title: "BIOL3100_Final_Project"
author: "Jessica Richards"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, echo=FALSE, include=FALSE}
library(tidyverse)
library(easystats)

```

# Tree Seedling Survival Responses Data

### Source

[Original Dataset](./Tree_Data.csv)

Link to original authors:
https://doi.org/10.5061/dryad.xd2547dpw 

Citation:
Wood, K., Kobe, R., Ibáñez, I., *et al.* (2023). Tree seedling functional traits
mediate plant-soil feedback survival responses across a gradient of light
availability. Dryad. https://doi.org/10.5061/dryad.xd2547dpw

Funded by: National Science Foundation: NSF DEB 145732, Michigan State University,
and Alma College

### About the data

Methods:

This study was a factorial blocked design field experiment. 4 tree species (*Acer saccharum, Prunus serotina, Quercus alba,* and *Quercus rubra*), 7 soil sources (*Acer saccharum, Prunus serotina, Quercus alba, Quercus rubra, Acer rebrum, Populus grandidentata*, and Sterilized) and a variety of forest understory light levels were used. There were a total of 3,024 seedlings in this experiment. Survival was measured twice a week and at 3 weeks, some seedlings were randomly selected to measure mycorrhizal colonization, phenolics, lignin, and NSC. 


Column information:

- No: Seedling ID number
- Plot: Plot the seedling was planted in (1-18)
- Subplot: Subplot seedling was planted in (A-E, one per corner and one in the middle)
- Species: Species of seedling (*Acer saccharum, Prunus serotina, Quercus alba*, and *Quercus rubra*)
- Light_ISF: Light level based on HemiView software.
- Light_Cat: Light levels were categorized into 'Low' 'Med' or 'High'
- Core: Year the soil core was taken, either 2016 or 2017
- Soil: Which species the soil core was taken from. Includes all 4 species as well as *Acer rubrum, Populus grandidentata* and a sterilized conspecific of each species
- Adult: Indicates which adult tree the soil core came from, up to 6 adults per species. 
- Sterile: Indicates whether the soil was steralized or not
- Conspecific: Indicates if the soil core used was from it's own species (conspecific), a different species (heterospecific) or steralized (sterile)
- Myco: Mycorrhizal type of the seedling (AMF or EMF)
- SoilMyco: Mycorrhizal type of the soil (AMF, EMF or Sterile)
- PlantDate: Date each tree was planted 
- AMF: Measures the percent arbuscular mycorrhizal fungi colonization on fine roots of harvested seedlings
- EMF: Measures the percent ectomycorrhizal fungi colonization on the root tips of harvested seedlings 
- Phenolics: Calculated as nmol Gallic acid equivalents per mg dry extract
- NSC: Calculated as the percent dry mass non-structural carbohydrates
- Lignin: Measured as percent dry mass lignin
- Census: The census number when each seedling died or was harvested
- Time: The number of days before the seedling died or was harvested 
- Event: Indicates whether the seedling died during the study (1 = dead, 0 = harvested or alive at end of study)
- Harvest: Indicates whether the seedling was harvested during the study
- Alive: Indicates whether the seedling was alive at the end of the study


Focus for Data Analysis:

The focus for this data analysis will be on how different environmental factors affect the amount of phenolics, lignin, and NSC (non-structural carbohydrates) in seedlings of different tree species. The main factors that will be analyzed are Light_Cat, Soil, Conspecific, Myco, SoilMyco, AMF, and EMF. 

## Original Data {.tabset}


### Output

Preview of Original Data

```{r original data output, echo=FALSE, include=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
glimpse(tree_dat)

```

### R code

Preview of Original Data

```{r original data r code, echo=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
glimpse(tree_dat)

```


# Cleaned Data

## Cleaned Data {.tabset}

### Output

The cleaned data has less columns and is better for plotting/modeling.

- Columns removed: Light_ISF, Core, Sterile, Adult, PlantDate, Census

- Columns 'Event', 'Harvest', and 'Alive' merged into one column called 'Event'
that indicates whether a tree died, was harvested, or was alive at the end of 
the study.
```{r clean data head, echo=FALSE, include=TRUE}
tree_dat_clean <- tree_dat %>%
  select(-c(Light_ISF, Core, Sterile, Adult, PlantDate, Census)) %>%
  mutate(Event = case_when(
    Event == '1' ~ 'X',
    Event == '0' ~ NA_character_)) %>%
  pivot_longer(cols = c(Event, Harvest, Alive),
               names_to = 'Event',
               values_to = 'Status') %>%
  mutate(Event = case_when(
    Event == 'Event' ~ 'Dead',
    TRUE ~ Event  )) %>%
  drop_na(Status) %>%
  select(-Status)

glimpse(tree_dat_clean)

```

### R code

The cleaned data has less columns and is better for plotting/modeling.

- Columns removed: Light_ISF, Core, Sterile, Adult, PlantDate, Census

- Columns 'Event', 'Harvest', and 'Alive' merged into one column called 'Event'
that indicates whether a tree died, was harvested, or was alive at the end of 
the study
```{r clean data r code, echo=TRUE}
tree_dat_clean <- tree_dat %>%
  select(-c(Light_ISF, Core, Sterile, Adult, PlantDate, Census)) %>%
  mutate(Event = case_when(
    Event == '1' ~ 'X',
    Event == '0' ~ NA_character_)) %>%
  pivot_longer(cols = c(Event, Harvest, Alive),
               names_to = 'Event',
               values_to = 'Status') %>%
  mutate(Event = case_when(
    Event == 'Event' ~ 'Dead',
    TRUE ~ Event  )) %>%
  drop_na(Status) %>%
  select(-Status)

glimpse(tree_dat_clean)

```


# Plotting Data



## Graph of NSC against Lignin {.tabset}


### Output

Graph of NSC v Lignin with Light_Cat as the color. 

NSC and Lignin have a positive relationship in *Acer saccharum* and *Prunus serotina*, and it seems like higher light levels result in higher Lignin and NSC levels whereas lower light levels result in lower Lignin and NSC levels. 
*Quercus alba* and *Quercus rubra* have a negative relationship between NSC and Lignin. 

```{r NSC v Lignin output, echo=FALSE, include=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat)) +
  geom_point() +
  facet_wrap(~Species, scales = 'free')

tree_dat_clean %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat)) +
  geom_smooth() +
  facet_wrap(~Species, scales = 'free')

```

### R code

Graph of NSC v Lignin

```{r NSC v Lignin r code, echo=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat)) +
  geom_point() +
  facet_wrap(~Species, scales = 'free')

tree_dat_clean %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat)) +
  geom_smooth() +
  facet_wrap(~Species, scales = 'free')

```

## AMF v Phenolics {.tabset}


### Output

Phenolics and AMF have a positive relationship in Prunus serotina, also plants
in high light category had more AMF and Phenolics levels 


```{r AMF v Phenolics output, echo=FALSE, include=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = AMF,
             y = Phenolics,
             color = Light_Cat)) +
  geom_point() +
  facet_wrap(~Species, scales = 'free')

```

### R code

Phenolics and AMF have a positive relationship in Prunus serotina, also plants
in high light category had more AMF and Phenolics levels 

```{r AMF v Phenolics r code, echo=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = AMF,
             y = Phenolics,
             color = Light_Cat)) +
  geom_point() +
  facet_wrap(~Species, scales = 'free')

```

## Phenolics  {.tabset}


### Output

Both SoilMyco and Conspecific affect Phenolics for at least one species

```{r Phenolics v SoilMyco and Conspecific output, echo=FALSE, include=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = SoilMyco,
             y = Phenolics,
             color = SoilMyco)) +
  geom_boxplot() +
  facet_wrap(~Species)


tree_dat_clean %>%
  ggplot(aes(x = Conspecific,
             y = Phenolics,
             color = Conspecific)) +
  geom_boxplot() +
  facet_wrap(~Species)

tree_dat_clean %>%
  ggplot(aes(x = Soil,
             y = Phenolics,
             color = Soil)) +
  geom_boxplot() +
  facet_wrap(~Species, scales = 'free')
```

### R code

Both SoilMyco and Conspecific affect Phenolics for at least one species

```{r Phenolics v SoilMyco and Conspecific r code, echo=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = SoilMyco,
             y = Phenolics,
             color = SoilMyco)) +
  geom_boxplot() +
  facet_wrap(~Species)


tree_dat_clean %>%
  ggplot(aes(x = Conspecific,
             y = Phenolics,
             color = Conspecific)) +
  geom_boxplot() +
  facet_wrap(~Species)

tree_dat_clean %>%
  ggplot(aes(x = Soil,
             y = Phenolics,
             color = Soil)) +
  geom_boxplot() +
  facet_wrap(~Species, scales = 'free')
```

## Light_Cat and Phenolics, Lignin, and NSC {.tabset}


### Output

Light category

```{r Light_Cat output, echo=FALSE, include=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = Light_Cat,
             y = Phenolics,
             color = Light_Cat)) +
  geom_jitter() +
  facet_wrap(~Species, scales = 'free')

tree_dat_clean %>%
  ggplot(aes(x = Light_Cat,
             y = Lignin,
             color = Light_Cat)) +
  geom_jitter() +
  facet_wrap(~Species, scales = 'free')

tree_dat_clean %>%
  ggplot(aes(x = Light_Cat,
             y = NSC,
             color = Light_Cat)) +
  geom_jitter() +
  facet_wrap(~Species, scales = 'free')

```

### R code

Light category
```{r Light_Cat r code, echo=TRUE}
tree_dat_clean %>%
  ggplot(aes(x = Light_Cat,
             y = Phenolics,
             color = Light_Cat)) +
  geom_jitter() +
  facet_wrap(~Species, scales = 'free')

tree_dat_clean %>%
  ggplot(aes(x = Light_Cat,
             y = Lignin,
             color = Light_Cat)) +
  geom_jitter() +
  facet_wrap(~Species, scales = 'free')

tree_dat_clean %>%
  ggplot(aes(x = Light_Cat,
             y = NSC,
             color = Light_Cat)) +
  geom_jitter() +
  facet_wrap(~Species, scales = 'free')

```



# Modeling Data


## Choosing a Model {.tabset}


### Output

The original plan was to make a model for predicting phenolics, a model for prediction lignin, and a model for predicting NSC, but the models for lignin and NSC had very high AIC values. This means that they likely aren't a good fit for the data and won't be very good at predicting values. 

The model for phenolics has a very high R-squared and a relatively low AIC meaning that it is a relatively good model. It predicts phenolics based on the interactions between Conspecific, Species, SoilMyco, AMF and Light_Cat. 

```{r choosing model output, echo=FALSE, include=TRUE}
#model for phenolics
mod1 <- glm(data = tree_dat_clean, Phenolics ~ Conspecific * Species * SoilMyco * AMF * Light_Cat)
mod2 <- glm(data = tree_dat_clean, Phenolics ~ Conspecific * Species * SoilMyco * Light_Cat)
mod3 <- glm(data = tree_dat_clean, Phenolics ~ Conspecific * Species * Light_Cat)

compare_performance(mod1, mod2, mod3)

#model for Lignin 
mod4 <- glm(data = tree_dat_clean, Lignin ~ Conspecific * Species * SoilMyco * Light_Cat)
mod5 <- glm(data = tree_dat_clean, Lignin ~ Species * SoilMyco * Light_Cat)
mod6 <- glm(data = tree_dat_clean, Lignin ~ Conspecific * Species * Light_Cat)

compare_performance(mod4, mod5, mod6)

#model for NSC
mod7 <- glm(data = tree_dat_clean, NSC ~ Conspecific * Species * SoilMyco * Light_Cat)
mod8 <- glm(data = tree_dat_clean, NSC ~ Species * SoilMyco * Light_Cat)
mod9 <- glm(data = tree_dat_clean, NSC ~ Conspecific * Species * Light_Cat)
mod10 <- glm(data = tree_dat_clean, NSC ~ Conspecific + Species + Light_Cat)

compare_performance(mod7, mod8, mod9, mod10)


```

### R code

The original plan was to make a model for predicting phenolics, a model for prediction lignin, and a model for predicting NSC, but the models for lignin and NSC had very high AIC values. This means that they likely aren't a good fit for the data and won't be very good at predicting values. 

The model for phenolics has a very high R-squared and a relatively low AIC meaning that it is a relatively good model. It predicts phenolics based on the interactions between Conspecific, Species, SoilMyco, AMF and Light_Cat. 


```{r choosing model r code, echo=TRUE}
#model for phenolics
mod1 <- glm(data = tree_dat_clean, Phenolics ~ Conspecific * Species * SoilMyco * AMF * Light_Cat)
mod2 <- glm(data = tree_dat_clean, Phenolics ~ Conspecific * Species * SoilMyco * Light_Cat)
mod3 <- glm(data = tree_dat_clean, Phenolics ~ Conspecific * Species * Light_Cat)

compare_performance(mod1, mod2, mod3)

#model for Lignin 
mod4 <- glm(data = tree_dat_clean, Lignin ~ Conspecific * Species * SoilMyco * Light_Cat)
mod5 <- glm(data = tree_dat_clean, Lignin ~ Species * SoilMyco * Light_Cat)
mod6 <- glm(data = tree_dat_clean, Lignin ~ Conspecific * Species * Light_Cat)

compare_performance(mod4, mod5, mod6)

#model for NSC
mod7 <- glm(data = tree_dat_clean, NSC ~ Conspecific * Species * SoilMyco * Light_Cat)
mod8 <- glm(data = tree_dat_clean, NSC ~ Species * SoilMyco * Light_Cat)
mod9 <- glm(data = tree_dat_clean, NSC ~ Conspecific * Species * Light_Cat)
mod10 <- glm(data = tree_dat_clean, NSC ~ Conspecific + Species + Light_Cat)

compare_performance(mod7, mod8, mod9, mod10)


```





# Conclusions
