---
title: "BIOL3100_Final_Project"
author: "Jessica Richards"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This is my data about trees

link to original authors:
https://doi.org/10.5061/dryad.xd2547dpw 

citation:
Wood, K., Kobe, R., Ibáñez, I., et al. (2023). Tree seedling functional traits
mediate plant-soil feedback survival responses across a gradient of light
availability. Dryad. https://doi.org/10.5061/dryad.xd2547dpw

funded by: National Science Foundation: NSF DEB 145732, Michigan State University,
and Alma College

### methods
This experiment was a factorial blocked design field experiment. 4 tree species 7 soil sources and a variety of forest understory light levels were used. There were a total of 3,024 seedlings. Survival was measured twice a week and seedlings were randomly selected each time to measure mycorrhizal colonization, phenolics, lignin, and NSC at 3 weeks. Cox proportional hazards survival models were used to evaluate survival and linear mixed effects models were used to test how light availability and soil source influence traits. 

okay but say that in your own words lol

### things to mention about data:

- generally what the study was
- where it came from (cite it)
- light ISF vs light Cat
- what core, soil, adult, conspecific mean
- Myco vs SoilMyco
-AMF ad EMF 
- Phenolics, NSC, Lignin
- Census vs Time
- Event, Harvest, and Alive, I want to make a column with all the trees, and then "dead, alive, or harvested" as the three options

## Results {.tabset}

### Plots
```{r tree data, echo=FALSE}
library(tidyverse)
tree_dat <- read.csv('Tree_Data.csv')


```

### Tables

We show the data in this tab.

```{r}
head(mtcars)
```


# This is me plotting the data 


## Results {.tabset}

### Plots

We show a scatter plot in this section.

```{r, fig.dim=c(5, 3)}
par(mar = c(4, 4, .5, .1))
plot(mpg ~ hp, data = mtcars, pch = 19)
```

### Tables

We show the data in this tab.

```{r}
head(mtcars)
```


