---
title: "BIOL3100_Final_Project"
author: "Jessica Richards"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, echo=FALSE, include=FALSE}
library(tidyverse)
library(easystats)

```

# This is my data about trees

### Source

[Original Dataset](./Tree_Data.csv)

Link to original authors:
https://doi.org/10.5061/dryad.xd2547dpw 

Citation:
Wood, K., Kobe, R., Ibáñez, I., et al. (2023). Tree seedling functional traits
mediate plant-soil feedback survival responses across a gradient of light
availability. Dryad. https://doi.org/10.5061/dryad.xd2547dpw

Funded by: National Science Foundation: NSF DEB 145732, Michigan State University,
and Alma College

### Methods/What this Data is
This experiment was a factorial blocked design field experiment. 4 tree species 7 soil sources and a variety of forest understory light levels were used. There were a total of 3,024 seedlings. Survival was measured twice a week and seedlings were randomly selected each time to measure mycorrhizal colonization, phenolics, lignin, and NSC at 3 weeks. Cox proportional hazards survival models were used to evaluate survival and linear mixed effects models were used to test how light availability and soil source influence traits. 

Column information

- No: Seedling ID number
- Plot: Plot the seedling was planted in (1-18)
- Subplot: Subplot seedling was planted in (A-E, one per corner and one in the middle)
- Species: Species of seedling (Acer saccharum, Prunus serotina, Quercus alba, and Quercus rubra)
- Light_ISF: Light level based on HemiView software.
- Light_Cat: Light levels were categorized into 'Low' 'Med' or 'High'
- Core: Year the soil core was taken, either 2016 or 2017
- Soil: Which species the soil core was taken from. Includes all 4 species as well as Acer rubrum, Populus grandidentata and a sterilized conspecific of each species
- Adult: Indicates which adult tree the soil core came from, up to 6 adults per species. 
- Sterile: Indicates whether the soil was steralized or not
- Conspecific: Indicates if the soil core used was from it's own species (conspecific), a different species (heterospecific) or steralized (sterile)
- Myco: Mycorrhizal type of the seedling (AMF or EMF)
- SoilMyco: Mycorrhizal type of the soil (AMF, EMF or Sterile)
- PlantDate: Date each tree was planted 
- AMF: Measures the percent arbuscular mycorrhizal fungi colonization on fine roots of harvested seedlings
- EMF: Measures the percent ectomycorrhizal fungi colonization on the root tips of harvested seedlings 
- Phenolics: Calculated as nmol Gallic acid equivalents per mg dry extract
- NSC: Calculated as the percent dry mass nonstructural carbohydrates
- Lignin: Measured as percent dry mass lignin
- Census: The census number when each seedling died or was harvested
- Time: The number of days before the seedling died or was harvested 
- Event: Indicates whether the seedling died during the study (1 = dead, 0 = harvested or alive at end of study)
- Harvest: Indicates whether the seedling was harvested during the study
- Alive: Indicates whether the seedling was alive at the end of the study

## Original Data {.tabset}


### Output

Preview of Original Data

```{r original data output, echo=FALSE, include=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
glimpse(tree_dat)

```

### R code

Preview of Original Data

```{r original data r code, echo=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
glimpse(tree_dat)

```


# Cleaned Data

## Cleaned Data {.tabset}

### Output

The cleaned data has less columns and is better for plotting/modeling.

- Columns removed: Light_ISF, Core, Sterile, Adult, PlantDate, Census

- Columns 'Event', 'Harvest', and 'Alive' merged into one column called 'Event'
that indicates whether a tree died, was harvested, or was alive at the end of 
the study
```{r clean data head, echo=FALSE, include=TRUE}
tree_dat_clean <- tree_dat %>%
  select(-c(Light_ISF, Core, Sterile, Adult, PlantDate, Census)) %>%
  mutate(Event = case_when(
    Event == '1' ~ 'X',
    Event == '0' ~ NA_character_)) %>%
  pivot_longer(cols = c(Event, Harvest, Alive),
               names_to = 'Event',
               values_to = 'Status') %>%
  mutate(Event = case_when(
    Event == 'Event' ~ 'Dead',
    TRUE ~ Event  )) %>%
  drop_na(Status) %>%
  select(-Status)

glimpse(tree_dat_clean)

```

### R code

The cleaned data has less columns and is better for plotting/modeling.

- Columns removed: Light_ISF, Core, Sterile, Adult, PlantDate, Census

- Columns 'Event', 'Harvest', and 'Alive' merged into one column called 'Event'
that indicates whether a tree died, was harvested, or was alive at the end of 
the study
```{r clean data r code, echo=TRUE}
tree_dat_clean <- tree_dat %>%
  select(-c(Light_ISF, Core, Sterile, Adult, PlantDate, Census)) %>%
  mutate(Event = case_when(
    Event == '1' ~ 'X',
    Event == '0' ~ NA_character_)) %>%
  pivot_longer(cols = c(Event, Harvest, Alive),
               names_to = 'Event',
               values_to = 'Status') %>%
  mutate(Event = case_when(
    Event == 'Event' ~ 'Dead',
    TRUE ~ Event  )) %>%
  drop_na(Status) %>%
  select(-Status)

glimpse(tree_dat_clean)

```


# Plotting Data





## Survival Subset {.tabset}


### Output

I created a subset of the cleaned data that includes a column of only seedlings 
that died or were alive at the end of the study. 
```{r survival subset, echo=FALSE, include=TRUE}
tree_dat_survival <- tree_dat_clean %>%
  mutate(Survival = Event) %>%
  mutate(Survival = case_when(Survival == 'Harvest' ~ NA_character_,
                   TRUE ~ Survival))

```

### R code

Preview of Original Data

```{r survival subset r code, echo=TRUE}
tree_dat_survival <- tree_dat_clean %>%
  mutate(Survival = Event) %>%
  mutate(Survival = case_when(Survival == 'Harvest' ~ NA_character_,
                   TRUE ~ Survival))

```







## Graph of NSC against Lignin {.tabset}


### Output

Graph of NSC v Lignin with Light category as the color and Survival as the shape. 
```{r NSC v Lignin output, echo=FALSE, include=TRUE}
tree_dat_survival %>%
  filter(!is.na(Survival)) %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat,
             shape = Survival)) +
  geom_point() +
  facet_wrap(~Species, scales = 'free')

tree_dat_survival %>%
  filter(!is.na(Survival)) %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat)) +
  geom_smooth() +
  facet_wrap(~Species, scales = 'free')

```

### R code

Graph of NSC v Lignin

```{r NSC v Lignin r code, echo=TRUE}
tree_dat_survival %>%
  filter(!is.na(Survival)) %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat,
             shape = Survival)) +
  geom_point() +
  facet_wrap(~Species, scales = 'free')

tree_dat_survival %>%
  filter(!is.na(Survival)) %>%
  ggplot(aes(x = NSC,
             y = Lignin,
             color = Light_Cat)) +
  geom_smooth() +
  facet_wrap(~Species, scales = 'free')

```



# Modeling Data


## Choosing a Model {.tabset}


### Output

Choosing a model: model 8 was best

```{r choosing model output, echo=FALSE, include=TRUE}
mod1 <-  glm(data = tree_dat_survival, Time ~ Soil + Lignin + NSC + Species)
mod2 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics )
mod3 <- glm(data = tree_dat_survival, Time ~ Species + Conspecific * Light_Cat + Phenolics + AMF + SoilMyco)
mod4 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * Lignin)
mod5 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * AMF)
mod6 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * SoilMyco)
mod7 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * AMF * SoilMyco)
mod8 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * NSC * Lignin)

compare_performance(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8)
compare_performance(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8) %>% plot()

best_model <- mod8

```

### R code

Choosing a model: model 8 was best


```{r choosing model r code, echo=TRUE}

mod1 <-  glm(data = tree_dat_survival, Time ~ Soil + Lignin + NSC + Species)
mod2 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics )
mod3 <- glm(data = tree_dat_survival, Time ~ Species + Conspecific * Light_Cat + Phenolics + AMF + SoilMyco)
mod4 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * Lignin)
mod5 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * AMF)
mod6 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * SoilMyco)
mod7 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * AMF * SoilMyco)
mod8 <- glm(data = tree_dat_survival, Time ~ Species * Conspecific * Light_Cat * Phenolics * NSC * Lignin)


compare_performance(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8)
compare_performance(mod1, mod2, mod3, mod4, mod5, mod6, mod7, mod8) %>% plot()

best_model <- mod8

```





# Conclusions
