---
title: "BIOL3100_Final_Project"
author: "Jessica Richards"
date: "2024-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, echo=FALSE, include=FALSE}
library(tidyverse)

```

# This is my data about trees

### Source

[Original Dataset](./Tree_Data.csv)

Link to original authors:
https://doi.org/10.5061/dryad.xd2547dpw 

Citation:
Wood, K., Kobe, R., Ibáñez, I., et al. (2023). Tree seedling functional traits
mediate plant-soil feedback survival responses across a gradient of light
availability. Dryad. https://doi.org/10.5061/dryad.xd2547dpw

Funded by: National Science Foundation: NSF DEB 145732, Michigan State University,
and Alma College

### Methods/What this Data is
This experiment was a factorial blocked design field experiment. 4 tree species 7 soil sources and a variety of forest understory light levels were used. There were a total of 3,024 seedlings. Survival was measured twice a week and seedlings were randomly selected each time to measure mycorrhizal colonization, phenolics, lignin, and NSC at 3 weeks. Cox proportional hazards survival models were used to evaluate survival and linear mixed effects models were used to test how light availability and soil source influence traits. 

Column information

- No: identification number (total # seeds)
- Plot: plot tree was planted in (how many plots?)
- Subplot: subplot tree was planted in (how many subplots?)
- Species: tree species (list species)
- Light_ISF: light reading in ISF (range?)
- Light_Cat: light readings were categorized into 'Low' 'Med' or 'High'
- Core: year the soil core was taken, either 2016 or 2017
- Soil: indicates which of the 4 species the soil core was taken from
- Adult: indicates which adult tree the seed came from (how many adults?)
- Sterile: indicates whether the soil was steralized or not
- Conspecific: indicates if soil core each seedling was planted in was taken from it's own
species (conspecific), a different species (heterospecific) or steralized (sterile)
- Myco: Myco of the seedling AMF or EMF
- SoilMyco: Myco of the soil AMF, EMF or Sterile
- PlantDate: date each tree was planted (how many plant dates?)
- AMF: 
- EMF: 
- Phenolics:
- Lignin:
- NSC:
- Census:?
- Time: how many days each tree survived
- Event: indicates whether tree died during the study
- Harvest: indicates whether tree was harvested during the study
- Alive: indicates whether tree was alive at the end of the study

## Original Data {.tabset}


### Output

Preview of Original Data

```{r original data output, echo=FALSE, include=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
glimpse(tree_dat)

```

### R code

Preview of Original Data

```{r original data r code, echo=TRUE}
tree_dat <- read.csv('Tree_Data.csv')
glimpse(tree_dat)

```


# Cleaned Data

### what do I want in cleaned data set 
- have only one Soil column with conspecific sterile, species, etc. 
- maybe combine soil columns?
- maybe combine myco columns?
- maybe combing AMF and EMF?
- maybe combine Phenolics, Lignin, and NSC

## Cleaned Data {.tabset}

### Output

The cleaned data has less columns and is better for plotting/modeling.

- Columns removed: Light_ISF, Core, Sterile, Adult, PlantDate, Census

- Columns 'Event', 'Harvest', and 'Alive' merged into one column called 'Event'
that indicates whether a tree died, was harvested, or was alive at the end of 
the study
```{r clean data head, echo=FALSE, include=TRUE}
tree_dat_clean <- tree_dat %>%
  select(-c(Light_ISF, Core, Sterile, Adult, PlantDate, Census)) %>%
  mutate(Event = case_when(
    Event == '1' ~ 'X',
    Event == '0' ~ NA_character_)) %>%
  pivot_longer(cols = c(Event, Harvest, Alive),
               names_to = 'Event',
               values_to = 'Status') %>%
  mutate(Event = case_when(
    Event == 'Event' ~ 'Dead',
    TRUE ~ Event  )) %>%
  drop_na(Status) %>%
  select(-Status)

glimpse(tree_dat_clean)

```

### R code

The cleaned data has less columns and is better for plotting/modeling.

- Columns removed: Light_ISF, Core, Sterile, Adult, PlantDate, Census

- Columns 'Event', 'Harvest', and 'Alive' merged into one column called 'Event'
that indicates whether a tree died, was harvested, or was alive at the end of 
the study
```{r clean data r code, echo=TRUE}
tree_dat_clean <- tree_dat %>%
  select(-c(Light_ISF, Core, Sterile, Adult, PlantDate, Census)) %>%
  mutate(Event = case_when(
    Event == '1' ~ 'X',
    Event == '0' ~ NA_character_)) %>%
  pivot_longer(cols = c(Event, Harvest, Alive),
               names_to = 'Event',
               values_to = 'Status') %>%
  mutate(Event = case_when(
    Event == 'Event' ~ 'Dead',
    TRUE ~ Event  )) %>%
  drop_na(Status) %>%
  select(-Status)

glimpse(tree_dat_clean)

```



# Plotting Data

# Modeling Data

# Conclusions
